`phyr`: An R package for phylogenetic species-distribution modelling in ecological communities

### 2019-09-27 ms.Rmd
pdf to png [skip ci]
Added my section on JSDMs

```{r designPlot, fig.cap="(ref:designPlotCap)", out.width="100%"}
knitr::include_graphics(normalize_path("designPlot.pdf"))
```

@@

pglmm is a joint species distribution model where the (residual) dependencies among species are modelled as a simple function of phylogeny [@wilkinson2019comparison].

>>>>

```{r designPlot, fig.cap="(ref:designPlotCap)", out.width="100%"}
knitr::include_graphics(normalize_path("designPlot.png"))
```

@@

Joint Species Distribution Models (JSDMs) are models where the response is some distribution of multiple species across sites or samples (measured as abundances or occurrences), where all species are modelled jointly, usually by allowing non-zero covariance between either species-level errors, species-level coefficients in the model, or both [@warton2015so]. A recent review and evaluation of JSDMs, for example, only considers models that allow modelling residual dependencies among species [@wilkinson2019comparison].

`pglmm` in `phyr` is a joint species distribution model where the (residual) dependencies among species are modelled as a simple function of phylogeny. JSDMs, and Species Distribution Models (SDM) in general, have typically been focused on producing accurate predictions of how species are distributed, usually in a geographic context. However, they can also be used for making inferences about the biology of species, such as which environmental factors are important in a species or set of species' biology, or whether traits or evolutionary history can help explain the structure of species communities. It is this kind of inference that is the focus of the JSDM implemented in `pglmm`. There is often a trade-off between improving predictions and making solid inferences, because increasing the complexity or flexibility of a model can improve its predictive power, but this same complexity makes it more difficult to understand what biology is being represented by the model outputs. In `pglmm`, the focus is on understanding whether evolutionary history is important in structuring species assemblages, which is facilitated by using a simply phylogenetic covariance model. This is in contrast to many other JSDMs [e.g. those in described in @wilkinson2019comparison], which attempt to flexibly estimate all pairwise species covariances, without taking into account any relationships amongst species, in an attempt to maximize predictive power.

Of course, the goals of solid inference and prediction are not mutually exclusive. Good prediction requires avoiding overfitting, which can be facilitated by reducing the number of parameters in the model. In some systems, it may be possible to make better predictions using a simple phylogenetic model if phylogeny is a strong predictor of species covariance, or if many species are poorly sampled and thus estimating covariances between them individually results in higher prediction variance. Using phylogeny can help closely related species share statistical strength through phylogenetic partial pooling. Ultimately, it can be very powerful to use the same kind of statistical framework to do both predictive and inferential work in ecology, because it allows ecologists to smoothly move between these two goals, both of which are highly important in modern ecology, and more easily and quickly draw mutual insights between them.
